---
title: "MRNGO pilot data analysis"
author: "Alex Ilyes"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

# Pre-requisits

Here we include options, import neccessary packages, and prepare plotting styles

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 9,
  fig.height = 6
)
```

```{r libraries}
library(tidyverse)
library(readxl)
library(knitr)
library(scales)
```

```{r plot-style}
poster_cols <- c(
  ink = "#111111",
  paper = "#DFE6F2",
  panel = "#F8F8F8",
  mauve = "#BFAFC9",
  blue = "#0072B2",
  green = "#009E73",
  sky = "#56B4E9",
  gold = "#E69F00",
  red = "#D55E00"
)

mrngo_theme <- function(base_size = 15) {
  theme_classic(base_size = base_size, base_family = "sans") +
    theme(
      plot.background = element_rect(fill = poster_cols["paper"], color = NA),
      panel.background = element_rect(fill = poster_cols["panel"], color = NA),
      panel.border = element_rect(fill = NA, color = poster_cols["ink"], linewidth = 0.5),
      axis.line = element_line(color = poster_cols["ink"], linewidth = 0.65),
      axis.ticks = element_line(color = poster_cols["ink"], linewidth = 0.65),
      axis.text = element_text(color = poster_cols["ink"]),
      axis.title = element_text(color = poster_cols["ink"], size = rel(1.08)),
      strip.background = element_rect(fill = poster_cols["mauve"], color = NA),
      strip.text = element_text(
        color = poster_cols["ink"],
        face = "bold",
        size = rel(1.02)
      ),
      legend.background = element_rect(
        fill = scales::alpha("white", 0),
        color = NA
      ),
      legend.key = element_rect(
        fill = scales::alpha("white", 0),
        color = NA
      )
    )
}

p_value_label <- function(p) {
  case_when(
    is.na(p) ~ "p = NA",
    p < .001 ~ "p < .001",
    TRUE ~ paste0("p = ", sub("^0", "", sprintf("%.3f", p)))
  )
}

block_labels <- c(run1 = "First block", run2 = "Repeat block")
measure_labels <- c(RT = "MIC latency (sec)", ACC = "Accuracy (% correct)")
block_cols <- c("First block" = "#0072B2", "Repeat block" = "#009E73")
participant_mauve <- "#BFAFC9"
participant_cols <- scales::hue_pal(l = 55, c = 80)

test_label <- function(test) {
  paste0("t(", round(unname(test$parameter), 1), ") = ", sprintf("%.2f", unname(test$statistic)), ", ", p_value_label(test$p.value))
}

cor_summary <- function(x, y) {
  d <- tibble(x = x, y = y) %>% drop_na()

  if (nrow(d) < 3 || sd(d$x) == 0 || sd(d$y) == 0) {
    return(tibble(n = nrow(d), r = NA_real_, p = NA_real_))
  }

  test <- cor.test(d$x, d$y, method = "pearson")

  tibble(
    n = nrow(d),
    r = unname(test$estimate),
    p = test$p.value
  )
}

partial_cor_summary <- function(x, y, z) {
  d <- tibble(x = x, y = y, z = z) %>% drop_na()

  if (nrow(d) < 4 || sd(d$x) == 0 || sd(d$y) == 0 || sd(d$z) == 0) {
    return(tibble(n_partial = nrow(d), partial_r = NA_real_, partial_p = NA_real_))
  }

  x_resid <- resid(lm(x ~ z, data = d))
  y_resid <- resid(lm(y ~ z, data = d))
  test <- cor.test(x_resid, y_resid, method = "pearson")

  tibble(
    n_partial = nrow(d),
    partial_r = unname(test$estimate),
    partial_p = test$p.value
  )
}

cor_label <- function(r, p, n) {
  if_else(
    is.na(r),
    paste0("n = ", n, ", ", p_value_label(p)),
    paste0("r = ", sprintf("%.2f", r), ", ", p_value_label(p), ", n = ", n)
  )
}
```

```{r figure-export-helper}
save_figure <- function(plot, name, width, height) {
  ggsave(file.path(figures_dir, paste0(name, ".png")), plot = plot, width = width, height = height, dpi = 300, bg = "white")
  ggsave(file.path(figures_dir, paste0(name, ".pdf")), plot = plot, width = width, height = height, device = cairo_pdf, bg = "white")
  svg_path <- file.path(figures_dir, paste0(name, ".svg"))
  tryCatch(
    ggsave(svg_path, plot = plot, width = width, height = height, bg = "white"),
    error = function(e) {
      grDevices::svg(svg_path, width = width, height = height, bg = "white")
      print(plot)
      grDevices::dev.off()
    }
  )
}
```

# READ-IN

We define paths for reading in preprocessed data.

```{r paths}

demography_path <- file.path("../derivatives", "mrngo_demography.xlsx")
mic_path <- file.path("../derivatives", "MIC_preprocessed.xlsx")
bar_path <- file.path("../derivatives", "BAR_preporcessed.xlsx")
figures_dir <- file.path("../derivatives", "figures/")

participants_summer <- c("P4647", "P7581", "P8504", "P4435", "P9581", "P3909")

```

MIC analyses include all participants (protocol unchanged). BAR analyses include only version 2 participants (due to change in task structure). 

```{r read-data}
demography_raw <- read_excel(demography_path)
mic_raw <- read_excel(mic_path)
bar_raw <- read_excel(bar_path)

demography <- demography_raw %>%
  transmute(
    ID = as.character(ID),
    version = as.numeric(version),
    sex = as.factor(sex),
    age_year = as.numeric(age_year),
    age_month = as.numeric(age_months),
    location = as.factor(location)
  )

version2_ids <- demography %>%
  filter(version == 2) %>%
  pull(ID)

version1_ids <- demography %>%
  filter(version == 1) %>%
  pull(ID)

version3_ids <- demography %>%
  filter(version == 3) %>%
  pull(ID)

mic <- mic_raw %>%
  mutate(
    ID = as.character(ID),
    run = as.integer(run),
    trialno = as.integer(trialno),
    conceptno = as.character(conceptno),
    concept = as.character(concept),
    concept_HU = as.character(concept_HU),
    rt = as.numeric(rt),
    correct = as.numeric(correct),
    responded = as.numeric(responded),
    run_label = paste0("run", dense_rank(run))
  )

bar <- bar_raw %>%
  filter(ID %in% participants_summer) %>%
  mutate(
    ID = as.character(ID),
    listno = as.integer(listno),
    trialno = as.integer(trialno),
    conceptno = as.character(conceptno),
    concept = as.character(concept),
    concept_HU = as.character(concept_HU),
    responded = as.numeric(responded),
    response = str_squish(str_to_lower(as.character(response))),
    similarity = as.numeric(similarity)
  )
```

# Demography Report

```{r demography-summary}
demography_overall <- demography %>%
  summarise(
    n = n(),
    mean_age_month = mean(age_month, na.rm = TRUE),
    sd_age_month = sd(age_month, na.rm = TRUE),
    mean_age_year = mean(age_year, na.rm = TRUE),
    sd_age_year = sd(age_year, na.rm = TRUE)
  )

sex_ratio_overall <- demography %>%
  count(sex, name = "n") %>%
  mutate(
    proportion = n / sum(n),
    percent = percent(proportion, accuracy = 0.1)
  )

demography_by_version <- demography %>%
  group_by(version) %>%
  summarise(
    n = n(),
    mean_age_month = mean(age_month, na.rm = TRUE),
    sd_age_month = sd(age_month, na.rm = TRUE),
    mean_age_year = mean(age_year, na.rm = TRUE),
    sd_age_year = sd(age_year, na.rm = TRUE),
    .groups = "drop"
  )

sex_ratio_by_version <- demography %>%
  count(version, sex, name = "n") %>%
  group_by(version) %>%
  mutate(
    proportion = n / sum(n),
    percent = percent(proportion, accuracy = 0.1)
  ) %>%
  ungroup()

kable(demography_overall, digits = 2)
kable(sex_ratio_overall)
kable(demography_by_version, digits = 2)
kable(sex_ratio_by_version)
```

# MIC Analysis

MIC trials are filtered for reaction-time outliers using a 3 SD rule within each participant and run. Accuracy summaries use the same filtered trial set so that RT and ACC are computed from matching trials.

```{r mic-outlier-filter}
mic_filtered <- mic %>%
  filter(!is.na(rt), !is.na(correct)) %>%
  group_by(ID, run_label) %>%
  mutate(
    participant_run_mean_rt = mean(rt, na.rm = TRUE),
    participant_run_sd_rt = sd(rt, na.rm = TRUE),
    rt_outlier = if_else(
      is.na(participant_run_sd_rt) | participant_run_sd_rt == 0,
      FALSE,
      abs(rt - participant_run_mean_rt) > 3 * participant_run_sd_rt
    )
  ) %>%
  ungroup() %>%
  filter(!rt_outlier)

mic_outlier_report <- mic %>%
  filter(!is.na(rt), !is.na(correct)) %>%
  left_join(
    mic_filtered %>%
      mutate(kept_after_rt_filter = TRUE) %>%
      select(ID, run_label, trialno, kept_after_rt_filter),
    by = c("ID", "run_label", "trialno")
  ) %>%
  mutate(kept_after_rt_filter = replace_na(kept_after_rt_filter, FALSE)) %>%
  group_by(ID, run_label) %>%
  summarise(
    trials_before_filter = n(),
    trials_after_filter = sum(kept_after_rt_filter),
    trials_removed = trials_before_filter - trials_after_filter,
    .groups = "drop"
  )

kable(mic_outlier_report)
```

## Run 1 and Run 2 RT and Accuracy

```{r mic-participant-run-summary}
mic_participant_run <- mic_filtered %>%
  group_by(ID, run_label) %>%
  summarise(
    mean_rt = mean(rt, na.rm = TRUE),
    accuracy = mean(correct, na.rm = TRUE),
    n_trials = n(),
    .groups = "drop"
  ) %>%
  left_join(demography %>% select(ID, version, age_month, sex), by = "ID")

mic_run_summary_long <- mic_participant_run %>%
  pivot_longer(
    cols = c(mean_rt, accuracy),
    names_to = "measure",
    values_to = "value"
  ) %>%
  mutate(
    measure = recode(measure, mean_rt = "RT", accuracy = "ACC"),
    measure = factor(measure, levels = c("RT", "ACC")),
    measure_label = factor(recode(as.character(measure), !!!measure_labels), levels = measure_labels),
    run_label = factor(run_label, levels = c("run1", "run2")),
    block_label = factor(recode(as.character(run_label), !!!block_labels), levels = block_labels),
    sex_label = factor(recode(as.character(sex), f = "Female", m = "Male", .default = as.character(sex))),
    plot_value = if_else(measure == "ACC", value * 100, value)
  )

kable(mic_participant_run, digits = 3)
```

```{r mic-paired-tests}
mic_run_wide <- mic_participant_run %>%
  select(ID, run_label, mean_rt, accuracy) %>%
  pivot_wider(
    names_from = run_label,
    values_from = c(mean_rt, accuracy)
  ) %>%
  drop_na(mean_rt_run1, mean_rt_run2, accuracy_run1, accuracy_run2)

mic_rt_ttest <- t.test(mic_run_wide$mean_rt_run1, mic_run_wide$mean_rt_run2, paired = TRUE)
mic_acc_ttest <- t.test(mic_run_wide$accuracy_run1, mic_run_wide$accuracy_run2, paired = TRUE)

mic_paired_tests <- tibble(
  measure = c("RT", "ACC"),
  t = c(unname(mic_rt_ttest$statistic), unname(mic_acc_ttest$statistic)),
  df = c(unname(mic_rt_ttest$parameter), unname(mic_acc_ttest$parameter)),
  p = c(mic_rt_ttest$p.value, mic_acc_ttest$p.value),
  label = c(test_label(mic_rt_ttest), test_label(mic_acc_ttest))
)

kable(mic_paired_tests, digits = 3)
```

```{r mic-run-barplots}
mic_plot_summary <- mic_run_summary_long %>%
  group_by(measure_label, block_label) %>%
  summarise(
    mean = mean(plot_value, na.rm = TRUE),
    se = sd(plot_value, na.rm = TRUE) / sqrt(sum(!is.na(plot_value))),
    .groups = "drop"
  )

mic_brackets <- mic_plot_summary %>%
  group_by(measure_label) %>%
  summarise(
    y = max(mean + se, na.rm = TRUE) * 1.08,
    .groups = "drop"
  ) %>%
  left_join(
    mic_paired_tests %>%
      mutate(measure_label = factor(recode(measure, !!!measure_labels), levels = measure_labels)) %>%
      select(measure_label, label),
    by = "measure_label"
  ) %>%
  mutate(x_start = 1, x_end = 2, x_mid = 1.5)

p_mic_barplots <- ggplot(mic_plot_summary, aes(x = block_label, y = mean, fill = block_label)) +
  geom_col(width = 0.62) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.08, linewidth = 1.05) +
  geom_point(
    data = mic_run_summary_long,
    aes(x = block_label, y = plot_value),
    inherit.aes = FALSE,
    position = position_jitter(width = 0.05, height = 0),
    size = 2.6,
    alpha = 0.8,
    color = participant_mauve
  ) +
  geom_segment(data = mic_brackets, aes(x = x_start, xend = x_end, y = y, yend = y), inherit.aes = FALSE, color = poster_cols["red"], linewidth = 1.05) +
  geom_segment(data = mic_brackets, aes(x = x_start, xend = x_start, y = y * 0.99, yend = y), inherit.aes = FALSE, color = poster_cols["red"], linewidth = 1.05) +
  geom_segment(data = mic_brackets, aes(x = x_end, xend = x_end, y = y * 0.99, yend = y), inherit.aes = FALSE, color = poster_cols["red"], linewidth = 1.05) +
  geom_text(data = mic_brackets, aes(x = x_mid, y = y, label = label), inherit.aes = FALSE, vjust = -0.4, size = 4.5, color = poster_cols["red"], family = "Myriad Pro") +
  facet_wrap(~ measure_label, scales = "free_y") +
  scale_fill_manual(values = block_cols) +
  labs(x = NULL, y = NULL) +
  mrngo_theme(base_size = 15) +
  theme(legend.position = "none")

p_mic_barplots
```

## Fastest Correctly Recognized Items

```{r mic-fastest-correct-items}
mic_correct_item_rt <- mic_filtered %>%
  filter(correct == 1, !is.na(rt)) %>%
  group_by(concept, concept_HU) %>%
  summarise(
    mean_correct_rt = mean(rt, na.rm = TRUE),
    sd_correct_rt = sd(rt, na.rm = TRUE),
    n_correct = n(),
    .groups = "drop"
  )

mic_fastest_correct_top5 <- mic_correct_item_rt %>%
  slice_min(mean_correct_rt, n = 5, with_ties = FALSE) %>%
  mutate(rank_type = "Fastest")

mic_fastest_correct_bottom5 <- mic_correct_item_rt %>%
  slice_max(mean_correct_rt, n = 5, with_ties = FALSE) %>%
  mutate(rank_type = "Slowest")

bind_rows(mic_fastest_correct_top5, mic_fastest_correct_bottom5) %>%
  arrange(rank_type, mean_correct_rt) %>%
  kable(digits = 3)
```

```{r mic-age-correlations}
mic_correlations <- mic_run_summary_long %>%
  filter(!is.na(age_month), !is.na(plot_value)) %>%
  group_by(measure_label, block_label) %>%
  summarise(cor_summary(age_month, plot_value), .groups = "drop") %>%
  mutate(label = cor_label(r, p, n))

mic_correlation_lines <- mic_run_summary_long %>%
  filter(!is.na(age_month), !is.na(plot_value)) %>%
  group_by(measure_label, block_label) %>%
  group_modify(~ {
    model <- lm(plot_value ~ age_month, data = .x)
    x_vals <- range(.x$age_month, na.rm = TRUE)
    tibble(
      x = x_vals[1],
      xend = x_vals[2],
      y = predict(model, newdata = tibble(age_month = x_vals[1])),
      yend = predict(model, newdata = tibble(age_month = x_vals[2]))
    )
  }) %>%
  ungroup()

plot_mic_correlation_block <- function(block_name) {
  plot_data <- mic_run_summary_long %>% filter(block_label == block_name)
  plot_lines <- mic_correlation_lines %>% filter(block_label == block_name)
  plot_labels <- mic_correlations %>% filter(block_label == block_name)

  ggplot(plot_data, aes(x = age_month, y = plot_value)) +
    geom_point(size = 3.0, alpha = 0.82, color = participant_mauve) +
    geom_segment(
      data = plot_lines,
      aes(x = x, xend = xend, y = y, yend = yend),
      inherit.aes = FALSE,
      color = poster_cols["ink"],
      linewidth = 1.1
    ) +
    geom_text(
      data = plot_labels,
      aes(x = -Inf, y = Inf, label = label),
      inherit.aes = FALSE,
      hjust = -0.02,
      vjust = 1.12,
      size = 4.4,
      family = "Myriad Pro"
    ) +
    facet_wrap(~ measure_label, scales = "free_y") +
    labs(x = "Age (months)", y = NULL) +
    mrngo_theme(base_size = 15) +
    theme(legend.position = "none")
}

p_mic_correlations_run1 <- plot_mic_correlation_block("First block")
p_mic_correlations_run2 <- plot_mic_correlation_block("Repeat block")

p_mic_correlations_run1
save_figure(p_mic_correlations_run1, "Figure1_MIC_correlations", width = 9, height = 5.2)

p_mic_correlations_run2

kable(mic_correlations %>% select(measure_label, block_label, n, r, p), digits = 3)
```

# BAR Analysis

BAR analyses use version 2 participants only.

## Guesses and Similarity Across Reveal Trials

Bars show the average number of guesses across participants and lists. Lines show participant-level average response similarity by concept and trial number; exact responses are highlighted in gold.

```{r bar-guesses-by-trial, fig.width=13, fig.height=6.5}
bar_trial_list <- bar %>%
  filter(trialno != 1) %>%
  group_by(ID, listno, concept, trialno) %>%
  summarise(
    list_guesses = sum(responded == 1, na.rm = TRUE),
    list_similarity = mean(similarity[responded == 1], na.rm = TRUE),
    exact_response = any(similarity == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(list_similarity = if_else(is.nan(list_similarity), NA_real_, list_similarity))

bar_trial_participant <- bar_trial_list %>%
  group_by(ID, concept, trialno) %>%
  summarise(
    participant_mean_guesses = mean(list_guesses, na.rm = TRUE),
    participant_mean_similarity = mean(list_similarity, na.rm = TRUE),
    exact_response = any(exact_response, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(participant_mean_similarity = if_else(is.nan(participant_mean_similarity), NA_real_, participant_mean_similarity)) %>%
  filter(!is.na(participant_mean_similarity)) %>%
  mutate(participant_mean_similarity = if_else(
    is.na(participant_mean_similarity),
    NA_real_,
    pmax(participant_mean_similarity, 0)
  ))

bar_trial_summary <- bar_trial_list %>%
  group_by(concept, trialno) %>%
  summarise(
    mean_guesses = mean(list_guesses, na.rm = TRUE),
    se_guesses = sd(list_guesses, na.rm = TRUE) / sqrt(sum(!is.na(list_guesses))),
    .groups = "drop"
  )

p_bar_guesses_similarity <- ggplot(bar_trial_summary, aes(x = trialno, y = mean_guesses)) +
  geom_col(fill = poster_cols["mauve"], width = 0.66) +
  geom_errorbar(aes(ymin = mean_guesses - se_guesses, ymax = mean_guesses + se_guesses), width = 0.08, linewidth = 1.0) +
  geom_line(
    data = bar_trial_participant,
    aes(x = trialno, y = participant_mean_similarity, color = ID, group = ID),
    inherit.aes = FALSE,
    linewidth = 0.9,
    alpha = 0.42,
    na.rm = TRUE
  ) +
  geom_point(
    data = bar_trial_participant %>% filter(!exact_response, !is.na(participant_mean_similarity)),
    aes(x = trialno, y = participant_mean_similarity, color = ID),
    inherit.aes = FALSE,
    size = 2.2,
    alpha = 0.72,
    na.rm = TRUE
  ) +
  geom_point(
    data = bar_trial_participant %>% filter(exact_response, !is.na(participant_mean_similarity)),
    aes(x = trialno, y = participant_mean_similarity),
    inherit.aes = FALSE,
    size = 3.8,
    color = poster_cols["gold"],
    alpha = 0.95,
    na.rm = TRUE
  ) +
  scale_x_continuous(breaks = sort(unique(bar_trial_summary$trialno))) +
  scale_color_manual(values = participant_cols(length(unique(bar_trial_participant$ID)))) +
  facet_wrap(~ concept, ncol = 5) +
  labs(x = "Trial number", y = "Average guesses / response similarity") +
  mrngo_theme(base_size = 15) +
  theme(legend.position = "none")

p_bar_guesses_similarity
save_figure(p_bar_guesses_similarity, "Figure2_BAR_guesses_similarity", width = 13, height = 6.5)

p_bar_guesses_similarity_B <- ggplot() +
  geom_line(
    data = bar_trial_participant,
    aes(x = trialno, y = participant_mean_similarity, color = ID, group = ID),
    linewidth = 0.9,
    alpha = 0.42,
    na.rm = TRUE
  ) +
  geom_point(
    data = bar_trial_participant %>% filter(!exact_response, !is.na(participant_mean_similarity)),
    aes(x = trialno, y = participant_mean_similarity, color = ID),
    size = 2.2,
    alpha = 0.72,
    na.rm = TRUE
  ) +
  geom_point(
    data = bar_trial_participant %>% filter(exact_response, !is.na(participant_mean_similarity)),
    aes(x = trialno, y = participant_mean_similarity),
    size = 3.8,
    color = poster_cols["gold"],
    alpha = 0.95,
    na.rm = TRUE
  ) +
  scale_x_continuous(breaks = sort(unique(bar_trial_summary$trialno))) +
  scale_color_manual(values = participant_cols(length(unique(bar_trial_participant$ID)))) +
  facet_wrap(~ concept, ncol = 5) +
  labs(x = "Trial number", y = "Response similarity") +
  mrngo_theme(base_size = 15) +
  theme(legend.position = "none")

p_bar_guesses_similarity_B
save_figure(p_bar_guesses_similarity_B, "Figure2B_BAR_guesses_similarity_no_bars", width = 13, height = 6.5)
```

## Response Frequency by Concept and Similarity

The `response` column is already in English, so responses are plotted directly from `bar`.

```{r bar-response-translation, fig.width=13, fig.height=6.5}
bar_response_words <- bar %>%
  filter(responded == 1, !is.na(response), response != "") %>%
  group_by(concept, response) %>%
  summarise(
    frequency = n(),
    mean_similarity = mean(similarity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    mean_similarity = replace_na(mean_similarity, 0),
    response_en = response
  ) %>%
  group_by(concept) %>%
  arrange(desc(frequency), response, .by_group = TRUE) %>%
  mutate(
    angle = 2 * pi * (row_number() - 1) / pmax(n(), 1),
    radius = 1 - pmin(pmax(mean_similarity, 0), 1),
    x = radius * cos(angle),
    y = radius * sin(angle)
  ) %>%
  ungroup()

bar_response_base <- ggplot(
  bar_response_words,
  aes(x = x, y = y, label = response_en, size = frequency, color = mean_similarity)
) +
  facet_wrap(~ concept, ncol = 5) +
  coord_equal(xlim = c(-1.18, 1.18), ylim = c(-1.18, 1.18)) +
  scale_size_continuous(range = c(3.0, 8.0)) +
  scale_color_gradient(low = poster_cols["blue"], high = poster_cols["red"], limits = c(0, 1)) +
  labs(x = NULL, y = NULL) +
  mrngo_theme(base_size = 15) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.line = element_blank(),
    legend.position = "none",
    strip.text = element_text(family = "Myriad Pro", face = "bold"),
    text = element_text(family = "Myriad Pro")
  )

p_bar_response_words <- if (requireNamespace("ggrepel", quietly = TRUE)) {
  bar_response_base +
    ggrepel::geom_text_repel(
      family = "Myriad Pro",
      fontface = "plain",
      max.overlaps = Inf,
      box.padding = 0.08,
      point.padding = 0.02,
      min.segment.length = Inf,
      seed = 1
    )
} else {
  bar_response_base +
    geom_text(
      family = "Myriad Pro",
      fontface = "plain",
      check_overlap = TRUE
    )
}

p_bar_response_words
save_figure(p_bar_response_words, "Figure3_BAR_response_similarity_words", width = 13, height = 6.5)
```

## BAR Participant Summaries

```{r bar-participant-summaries}
bar_participant_list <- bar %>%
  group_by(ID, listno) %>%
  summarise(
    concept = first(concept),
    response_number = sum(responded == 1, na.rm = TRUE),
    exact_answer = any(similarity == 1, na.rm = TRUE),
    mean_similarity = mean(similarity[responded == 1], na.rm = TRUE),
    first_exact_reveal = if_else(
      any(similarity == 1, na.rm = TRUE),
      min(trialno[similarity == 1], na.rm = TRUE) / 10,
      NA_real_
    ),
    .groups = "drop"
  ) %>%
  mutate(mean_similarity = if_else(is.nan(mean_similarity), NA_real_, mean_similarity))

bar_participant_summary <- bar_participant_list %>%
  group_by(ID) %>%
  summarise(
    first_exact_reveal = mean(first_exact_reveal, na.rm = TRUE),
    average_response_similarity = mean(mean_similarity, na.rm = TRUE),
    bar_accuracy = mean(exact_answer, na.rm = TRUE),
    average_response_number = mean(response_number, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(across(c(first_exact_reveal, average_response_similarity), ~ if_else(is.nan(.x), NA_real_, .x)))

kable(bar_participant_summary, digits = 3)

```

# BAR version comparison

```{r}
bar_compare <- bar_raw %>%
  mutate(
    ID = as.character(ID),
    listno = as.integer(listno),
    trialno = as.integer(trialno),
    concept = as.character(concept),
    responded = as.numeric(responded),
    similarity = as.numeric(similarity)
  ) %>%
  left_join(
    demography %>% select(ID, version),
    by = "ID"
  )

bar_trial <- bar_compare %>%
  filter(trialno != 1) %>%
  group_by(ID, version, listno, concept, trialno) %>%
  summarise(
    exact_now = any(responded == 1 & similarity == 1, na.rm = TRUE),
    guesses = sum(responded == 1, na.rm = TRUE),
    best_similarity = if_else(
      any(responded == 1 & !is.na(similarity)),
      max(similarity[responded == 1], na.rm = TRUE),
      NA_real_
    ),
    .groups = "drop"
  ) %>%
  group_by(ID, version, listno, concept) %>%
  arrange(trialno, .by_group = TRUE) %>%
  mutate(
    opportunity = row_number(),
    n_opportunities = n(),
    reveal_proportion = opportunity / n_opportunities,
    exact_by_now = as.logical(cummax(exact_now))
  ) %>%
  ungroup()
```


```{r}
bar_curve <- bar_trial %>%
  group_by(ID, version, reveal_proportion) %>%
  summarise(
    proportion_exact = mean(exact_by_now),
    .groups = "drop"
  )

bar_curve_summary <- bar_curve %>%
  group_by(version, reveal_proportion) %>%
  summarise(
    mean_exact = mean(proportion_exact),
    se = sd(proportion_exact) / sqrt(n()),
    .groups = "drop"
  )

ggplot(
  bar_curve_summary,
  aes(reveal_proportion, mean_exact, color = factor(version))
) +
  geom_ribbon(
    aes(
      ymin = pmax(mean_exact - 1.96 * se, 0),
      ymax = pmin(mean_exact + 1.96 * se, 1),
      fill = factor(version)
    ),
    alpha = .15,
    color = NA
  ) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.4) +
  scale_x_continuous(
    labels = percent_format(),
    limits = c(0, 1)
  ) +
  scale_y_continuous(
    labels = percent_format(),
    limits = c(0, 1))+
  mrngo_theme() +
  labs(
    x = "Proportion of trial opportunities used",
    y = "Concepts identified exactly",
    color = "BAR version",
    fill = "BAR version"
  ) 

save_figure(last_plot(), "Figure_BAR_learning_curve", width = 9, height = 5.2)
```

```{r}
bar_by_six <- bar_trial %>%
  filter(opportunity <= 6) %>%
  group_by(ID, version, listno, concept) %>%
  summarise(
    exact_by_six = any(exact_now),
    .groups = "drop"
  ) %>%
  group_by(ID, version) %>%
  summarise(
    accuracy_by_six = mean(exact_by_six),
    n_concepts = n(),
    .groups = "drop"
  )

ggplot(
  bar_by_six,
  aes(factor(version), accuracy_by_six, color = factor(version))
) +
  geom_boxplot(
    width = .35,
    outlier.shape = NA,
    color = poster_cols["ink"]
  ) +
  geom_jitter(width = .08, height = 0, size = 2.7, alpha = .75) +
  stat_summary(
    fun = mean,
    geom = "point",
    shape = 18,
    size = 4,
    color = poster_cols["ink"]
  ) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  labs(
    x = "BAR version",
    y = "Exact-answer rate after six opportunities"
  ) +
  mrngo_theme() +
  theme(legend.position = "none")
save_figure(last_plot(), "Figure_BAR_accuracy_by_six", width = 9, height = 5.2)

```

```{r}
bar_list_summary <- bar_trial %>%
  group_by(ID, version, listno, concept) %>%
  summarise(
    n_opportunities = first(n_opportunities),
    exact_answer = any(exact_now),
    first_exact_opportunity = if_else(
      exact_answer,
      min(opportunity[exact_now]),
      NA_integer_
    ),
    proportion_used = if_else(
      exact_answer,
      first_exact_opportunity / n_opportunities,
      NA_real_
    ),
    guess_rate = sum(guesses) / n_opportunities,
    .groups = "drop"
  )
kable(bar_list_summary %>% group_by(version) %>% summarise(
  mean_proportion_used = mean(proportion_used, na.rm = TRUE),
  sd_proportion_used = sd(proportion_used, na.rm = TRUE),
  mean_guess_rate = mean(guess_rate, na.rm = TRUE),
  sd_guess_rate = sd(guess_rate, na.rm = TRUE),
  .groups = "drop"
), digits = 3)
```

```{r}
plot_bar_proportion_used <- ggplot(
  bar_list_summary,
  aes(factor(version), proportion_used, color = factor(version))
) +
  geom_boxplot(
    width = .35,
    outlier.shape = NA,
    color = poster_cols["ink"]
  ) +
  geom_jitter(width = .08, height = 0, size = 2.7, alpha = .75) +
  stat_summary(
    fun = mean,
    geom = "point",
    shape = 18,
    size = 4,
    color = poster_cols["ink"]
  ) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  labs(
    x = "BAR version",
    y = "Proportion of trial opportunities used for first exact answer"
  ) +
  mrngo_theme() +
  theme(legend.position = "none")
plot_bar_proportion_used
```


# MIC vs BAR

```{r mic-bar-correlations, fig.width=11, fig.height=5}
mic_run1_summary <- mic_participant_run %>%
  filter(run_label == "run1") %>%
  select(ID, mic_latency = mean_rt, mic_acc = accuracy)

mic_bar_summary <- mic_run1_summary %>%
  inner_join(bar_participant_summary, by = "ID") %>%
  left_join(demography %>% select(ID, age_month), by = "ID") %>%
  filter(abs(mic_latency - mean(mic_latency, na.rm = TRUE)) <= 2 * sd(mic_latency, na.rm = TRUE)) %>%
  mutate(age_group = factor(
    if_else(age_month < 84, "Age below 7 years", "Age above 7 years"),
    levels = c("Age below 7 years", "Age above 7 years")
  ))

mic_bar_age_cols <- c(
  "Age below 7 years" = unname(poster_cols["blue"]),
  "Age above 7 years" = unname(poster_cols["green"])
)

mic_bar_summary %>%
  count(age_group) %>%
  kable()

mic_bar_long <- bind_rows(
  mic_bar_summary %>%
    transmute(
      ID,
      age_month,
      age_group,
      comparison = "CFC feature at first correct - MIC latency",
      x_label = "CFC feature at first correct",
      y_label = "MIC latency (sec)",
      x = first_exact_reveal * 10,
      y = mic_latency
    ),
  mic_bar_summary %>%
    transmute(
      ID,
      age_month,
      age_group,
      comparison = "CFC average similarity - MIC accuracy",
      x_label = "CFC average similarity",
      y_label = "MIC accuracy (% correct)",
      x = average_response_similarity,
      y = mic_acc * 100
    )
)

mic_bar_correlations <- mic_bar_long %>%
  group_by(comparison, x_label, y_label) %>%
  summarise(
    cor_summary(x, y),
    partial_cor_summary(x, y, age_month),
    .groups = "drop"
  ) %>%
  mutate(label = paste0(
    "r = ", sprintf("%.2f", r), ", ", p_value_label(p),
    "\npartial r = ", sprintf("%.2f", partial_r), ", ", p_value_label(partial_p)
  ))

plot_mic_bar_panel <- function(comparison_name) {
  plot_data <- mic_bar_long %>% filter(comparison == comparison_name)
  plot_label <- mic_bar_correlations %>% filter(comparison == comparison_name)

  ggplot(plot_data, aes(x = x, y = y)) +
    geom_point(aes(color = age_group), size = 3.2, alpha = 0.9) +
    geom_smooth(method = "lm", se = FALSE, color = poster_cols["ink"], linewidth = 1.15) +
    geom_text(
      data = plot_label,
      aes(x = -Inf, y = Inf, label = label),
      inherit.aes = FALSE,
      hjust = -0.02,
      vjust = 1.12,
      size = 4.1,
      family = "Myriad Pro"
    ) +
    scale_color_manual(values = mic_bar_age_cols, drop = FALSE, na.value = poster_cols["ink"]) +
    labs(x = first(plot_data$x_label), y = first(plot_data$y_label), color = NULL) +
    mrngo_theme(base_size = 15) +
    theme(
      plot.background = element_rect(fill = poster_cols["paper"], color = NA),
      panel.background = element_rect(fill = poster_cols["panel"], color = NA),
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.background = element_rect(fill = "white", color = poster_cols["ink"], linewidth = 0.4),
      legend.box.background = element_rect(fill = "white", color = poster_cols["ink"], linewidth = 0.4),
      legend.key = element_rect(fill = "white", color = NA)
    )
}

p_mic_bar_latency <- plot_mic_bar_panel("CFC feature at first correct - MIC latency")
p_mic_bar_accuracy <- plot_mic_bar_panel("CFC average similarity - MIC accuracy")

p_mic_bar_correlations <- if (requireNamespace("patchwork", quietly = TRUE)) {
  patchwork::wrap_plots(p_mic_bar_latency, p_mic_bar_accuracy, nrow = 1, guides = "collect") &
    theme(
      plot.background = element_rect(fill = poster_cols["paper"], color = NA),
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.background = element_rect(fill = "white", color = poster_cols["ink"], linewidth = 0.4),
      legend.box.background = element_rect(fill = "white", color = poster_cols["ink"], linewidth = 0.4)
    )
} else {
  p_mic_bar_latency
}

p_mic_bar_correlations
ggsave(file.path(figures_dir, "Figure5_MIC_BAR_correlations.png"), plot = p_mic_bar_correlations, width = 11, height = 5.8, dpi = 300, bg = poster_cols["paper"])
ggsave(file.path(figures_dir, "Figure5_MIC_BAR_correlations.pdf"), plot = p_mic_bar_correlations, width = 11, height = 5.8, device = cairo_pdf, bg = poster_cols["paper"])
tryCatch(
  ggsave(file.path(figures_dir, "Figure5_MIC_BAR_correlations.svg"), plot = p_mic_bar_correlations, width = 11, height = 5.8, bg = poster_cols["paper"]),
  error = function(e) {
    grDevices::svg(file.path(figures_dir, "Figure5_MIC_BAR_correlations.svg"), width = 11, height = 5.8, bg = poster_cols["paper"])
    print(p_mic_bar_correlations)
    grDevices::dev.off()
  }
)

kable(mic_bar_correlations %>% select(comparison, n, r, p, n_partial, partial_r, partial_p), digits = 3)
```
